\textbf{Computational Fuzzy Extractors}

### 2013-03-07 CompFuzzExt.tex
thm 5.3 cleaned...

%This problem can be stated in the language of indistinguishability using the Decision-to-search reduction of Regev or Peikert:
\begin{lemma}~\protect{~\cite[Corollary 10]{dottling2012}}
\label{lem:uniform LWE decision}
Let $n$ be a security parameter.  Let $q = q(n) = \poly(n)$ be a prime integer and $m = m(n) = \poly(n)$ be an integer. Let $\gamma\in (0, 1)$ be a constant and let $\rho\in (0,1)$ such that $\rho q \geq 2n^{1/2+3\gamma}m$. Assume decision-version of the shortest vector problem (GAPSVP) and the shortest independent vectors problem (SIVP) are hard to approximate~(on lattices of dimension $n/2$) with the factor $\tilde{O}(n^{1/2+3\gamma}m/\rho)$ for quantum algorithms in the worst case, then, for $\chi = [-\rho q, \rho q]$, then $\distLWE_{n, m, q, \chi}$ is $(\epsilon, s_{sec})\mbox{-}$secure, where $\epsilon = \ngl(n)$ for $s_{sec}=\poly(n)$.
\end{lemma}

To extract pseudoranom bits, we use a result of Akavia, Goldwasser, and Vaikuntanathan~\cite{akavia2009} to show that $X$ has many hardcore bits.  The result says that if $\distLWE_{(n-k, m, q, \chi)}$ is secure then any $k$ variables of $X$ in a $\distLWE_{(n, m, q, \chi)}$ instance are hardcore.  We state their result for a general error distribution~(noting their proof does not consider the error distribution anywhere):
\begin{lemma}\protect{\cite[Lemma 2]{akavia2009}}
\label{lem:many hardcore bits}
Let $n>0$ be an integer, $q\geq 2$, let $\chi$ be an error distribution over $\zq$ and $k\leq n$ be an integer.  Let $\vA\in \zq^{m\times n} , X\in\zq^n$ and $E$ be distributed according to $\chi^m$.  If $\distLWE_{(n-k, m, q, \chi)}$ is $\epsilon, s_{sec}$ hard then $H^{\hill}_{\epsilon, s_{sec}'}(X_{1,..., k}| \vA, \vA X+E )\geq k\log q$ where $s_{sec}' \approx s_{sec} - n^3$.
\end{lemma}

@@

\section{\consref{cons:informal construction} with Nonuniform Sources}
\subsection{\consref{cons:informal construction} with Symbol Fixing Sources}
\label{sec:LWE block fixing sources}
In this section, we show the security of~\consref{cons:informal construction} is preserved for certain high entropy distributions.  Showing security of a computational fuzzy extractor for arbitrary high min-entropy distributions is an open problem.  First we recall the notion of a symbol fixing source~(from~\cite{KZ07}): 
\begin{definition}
Let $W = (W_1,..., W_{m+\alpha})$ be a distribution where $W_i$ takes values over $\{0,1\}^b$, we say that a distribution is a $m$-\emph{symbol fixing source} if $H_\infty(W)\geq mb$ and for all $W_i$ one of the following conditions is met:
\begin{itemize}
\item $W_i \overset{d}= U_b$
\item There exists a value $c_i$ such that $\Pr[W_i = c_i] =1$.
\end{itemize}
\end{definition}
%A block fixing source is an extension of a entropy source where each bit is independently random or fixed to larger alphabets.

@@

\begin{theorem}
\label{thm:lossless block sketch log}
Let $n$ be a security parameter and let $t = c\log n$ for some constant $c$.  Let $d\in\Fq^+$ be a constant and consider the Hamming metric with block length $b = \log (2n^3 (c+d)/d)$.  Let $W$ be $\alpha$-symbol fixing source over $\zo^{((c+d)nb/d+\alpha b}$.  Assume GAPSVP and SIVP are hard to approximate within polynomial factors~(on lattices of dimension $n/2$) using quantum algorithms.   There is a setting of $q = \poly(n)$ such that \consref{cons:informal construction} is a lossless fuzzy extractor.  That is, $H^{\hill}_{\epsilon, s_{sec}}(X_{1,..., n/2} |\vA, \vA X+W) = H_\infty(W)$ for %the following parameters:
%\[
%(\zo^{((c+d)n/d+\alpha)b}, \Hoo(W),\Hoo(W), \epsilon, s_{sec}, s_{sketch}, s_{rec}, t)\text{-computational secure sketch}
%\]
$\epsilon = \ngl(n), s_{sec} = \poly(n)$, sketch time $O(n^3\log n)$, and expected recover time $O(n^{2d+3} \log n)$.
\end{theorem}

\subsection{\consref{cons:informal construction} with slightly deficient sources}
\label{sec:DP general error LWE}
In this section, we discuss how to generalize \consref{cons:informal construction} to an arbitrary high min-entropy distribution.  We will show security using a result of Micciancio and Peikert~\cite{micciancio2013hardness}.  This result allows us to argue security for an arbitrary distribution that is nearly high entropy.  This result should be seen as a complement to the proceeding sections.  It allows for a more general error distribution, but with worse parameters than \thref{thm:lossless secure conductor log}.  Also \thref{thm:lossless block sketch log} allows for more deficient sources but of a more restricted form~(symbol fixing sources).
\begin{theorem}
Let $n$ be a security parameter and let $m=cn, \ell = m-n/2$.  Let $C$ be a constant and define the parameter $\rho \geq (Cm)^{\frac{m-n/2}{n/2}} = (Cm)^{2c-1}$ and let $q$ be a prime where $\poly(n) = q\geq (4\rho)^{\frac{c}{c-1}}$.  Consider a distribution $\chi$ over the space $\{-\rho, ..., \rho \}^m$.  If $\Hoo(\chi)\geq m\log \rho$ and SIVP is hard to approximate in the worst case for quantum algorithms within an approximation factor of $\gamma=\tilde{O}(\sqrt{n} q)$~(on lattices of $n/2$ dimensions), then $\distLWE_{n, m, q, \chi}$ is $(\epsilon, s_{sec})\mbox{-}$secure, where $\epsilon = \ngl(n)$ for $s_{sec}=\poly(n)$.
\end{theorem}
\textbf{Note:}  The above theorem is a generalized version of~\cite[Theorem 4.6]{micciancio2013hardness}.  In particular, we note the requirement of sampling from the uniform distribution over some set $\chi \subset \{-\rho,..., \rho\}^m$ is not necessary.  The only requirement in the proof is that $|\chi|\geq m\log \rho$.  Requiring that $\Hoo(\chi) = m\log \rho$ implies that the support of $\chi$ contains at least $2^{m\log \rho}$ elements.  This theorem allows all distributions over a space of size $(2\rho)^m$ that have entropy at least $m\log \rho$.  This means we support all distributions whose entropy gap is no more than $m$ bits.

\bnote{Need to continue and go through their Lemma 2.1 and 2.4 to be sure I believe this.  Everything else should go through fine.}

\bnote{We need to turn this hardness result into a statement about computational fuzzy extractors.  Use the Akavia result.  Then we need to suggest parameters.  How much detail should we go into?}

@@

\begin{claim}
\label{cl:randomMatrixDist}
The matrix $\vA$ generated after Step \ref{step:randomization} is a random matrix.
\end{claim}
\begin{proof}
In the randomized matrix $\vect{A}$, for $1\le i \le m$ and $n+1\le j \le n+\alpha$, $A_{ij}$ is assigned the value $\gamma_{i,(j - n)}$, which is a random value.  Similarly, for $m+1\le i \le m+\alpha$ and $n+1\le j \le n+\alpha$, each entry $A_{ij}$ is random due to the original construction of the matrix $\vA^*$.  Then before step \ref{step:randomization} the matrices $\vA', \vect{R}$ are random submatrices.  Thus, it remains to show that $\vA', \vect{R}$ remain random after step 5. For $1\le i \le m+\alpha$ and $1\le j \le n$, by construction $A_{ij} = A_{ij}'+\sum_{k=1}^\alpha \gamma_{i, k} R_{kj}$, which is a truly random number. Overall, each row vector $\va_{i}$ in $\vA$ is the sum of a random vector and a random linear combination independent vectors. Therefore, the entire matrix $\vA$ is a truly random matrix.  This completes the claim.
\end{proof}
\begin{claim}
\label{cl:random ax+e}
If $\D'$ is provided with input distributed as $\vA', \vA'\vx'+\ve'$ then $\vb = \vA \vx+\ve$ where $x_i = x_i'$ for $1\leq i \leq n$ and $x_{n+i} = b_{m+i}^*$ is uniformly generated otherwise and  $e_i = e_i'$ for $1\leq i\leq m$ and $e_i = 0$ for $m<i\leq m+\alpha$.
\end{claim}
\begin{proof}
Before step \ref{step:randomization} the claim is satisfied.  Denote by $\vx'$ the value such that $\vA'\vx'+\ve = \vb$.  
Then define $\vx^*$ as:
\[
x^*_i = \begin{cases}
x_i', & \text{if }1\leq i\leq n \\
b_{m-n+i}^*-\sum_{j=1}^n R_{i, j} x_j', & n+1\leq i\leq n+\alpha\,.
\end{cases}
\]
Then $\vx^*$ is uniformly random and $\vb^* = \vA^*\vx^* + \ve^*$~(where $\ve^*$ is $\ve'$ with $\alpha$ 0s appended).  Thus, it remains to show that step~\ref{step:randomization} preserves a solution.
%For convenience, denote by $\vx^*$ the vector where $x_i^* = x_i'$ for $1\leq i \leq n$ and $x_{n+i}^* = b_{m+i}^*$ otherwise.  
For a matrix $\vA$ we will denote the $i$-th row by $\va_i$~(similarly for the matrix formed by $\gamma_{i,j}$).  Thus it suffices to show for all rows $i$, if $b_i^* = \va^*_i \vx^*+e^*_i$  then $b_i = \va_i \vx^* + e^*_i$.
%, we have added $\alpha$ equations as well as $\alpha$ unknowns, furthermore, those $\alpha$ equations have no error.  The number of unknowns we added are $x_{n+i} = b^*_{m+i} - \sum^{n}_{j=1}R_{ij}x_j$ for $i=n+1,..., n+\alpha$.
\bnote{We need to justify these steps better.}
\bnote{Is it really standard to go to lower case for the rows?}
\xnote{lower case for vectors, which is row vector in the matrix}

\xnote{Needs to be clean up the $\gamma$, it's not correct, at least not consistent with the randomized algorithm.}
We have the following for all $i=1,..., m+\alpha$:
\begin{align*}
\va_i \vx^* + e_i^* &=\va^*_i\vx^*+\sum_{k=1}^\alpha \gamma_{i} \va^*_{m+k}\vx^*+e_i^*\\
&= \va_i^*\vx^* + e_i^* + \sum_{k=1}^\alpha \mathbf{\gamma}_{i} \va^*_{m+k}\vx^*\\
&= b_i^* + \sum_{k=1}^\alpha\gamma_i \va_{m+k}^* \vx^*\\
&= b_i^* + \sum_{k=1}^\alpha \gamma_i \left(\left( \sum_{j=1}^n \vR_{i, j}x_j^* \right) + x_{n+k}^*\right)\\
&= b_i^* +\sum_{k=1}^\alpha\gamma_i b_{m+k}^*= b_i
\end{align*}
%\begin{align*}
%\vA_i \vx + e_i &= \sum_{j=1}^{n+\alpha} A_{ij} x_j +e_i = \sum_{j=1}^{n+\alpha}(A^*_{ij}x_j+\sum_{k=1}^\alpha \gamma_{i, k} A^*_{(m+k)j}x_j)+e_i\\
%& = \sum_{j=1}^{n+\alpha}(A^*_{ij}x_j)+e_i +\sum_{k=1}^\alpha \gamma_{i, k} \sum_{j=1}^{n+\alpha} (A^*_{(m+k)j}x_j)\\
%& = b^*_i +  \sum_{k=1}^\alpha\gamma_{i, k} b^*_{m+k} = (\va_i^* \vx + e_i) + \sum_{k=1}^\alpha\gamma_{i, k} (\va_{m+k}^* \vx + e_i)\\
%& = b_i \label{reductionEq}
%\end{align*}
\end{proof}
\begin{claim} 
\label{clm:random b}
If $\D'$ is provided with the input distributed as $\vA', \vb'$ where $\vb'$ is uniformly random, then $\vb$ is uniformly random conditioned on $\vA$.
\end{claim}
\begin{proof}
Before step~\ref{step:randomization}, each $b_i^*$ is uniformly and independently random.  After step~\ref{step:randomization}, since the $b_i*$ and $\gamma_i$ are random, then each $b_i$ is a fresh random linear combination of random values:
\[
b_i = b_i^* + \sum_{k=1}^\alpha \gamma_i b_{m+k}^*.
\]
%We also need to show that $\vb$ is uniformly distributed even conditioned on $\vA$. Step~\ref{step:randomization} randomized each

$\vb$ doesn't dependent on $\vA$ simply because $\vb'$ doesn't dependent on $\vA'$ \xnote{I need show more work on this, maybe?}
Thus, each of these values is uniformly and independently random.
\end{proof}

>>>>

%This problem can be stated in the language of indistinguishability using the Decision-to-search reduction of Regev or Peikert:
\begin{lemma}~\protect{~\cite[Corollary 10]{dottling2012}}
\label{lem:uniform LWE decision}
Let $n$ be a security parameter.  Let $q = q(n) = \poly(n)$ be a prime integer and $m = m(n) = \poly(n)$ be an integer. Let $\gamma\in (0, 1)$ be a constant and let $\rho\in (0,1)$ such that $\rho q \geq 2n^{1/2+3\gamma}m$. Assume decision-version of the shortest vector problem (GAPSVP) and the shortest independent vectors problem (SIVP) are hard to approximate~(on lattices of dimension $n$) with the factor $\tilde{O}(n^{1/2+3\gamma}m/\rho)$ for quantum algorithms in the worst case, then, for $\chi = [-\rho q, \rho q]$, the $\distLWE_{n, m, q, \chi}$ is $(\epsilon, s_{sec})\mbox{-}$secure, where $\epsilon = \ngl(n)$ for $s_{sec}=\poly(n)$.
\end{lemma}

To extract pseudoranom bits, we use a result of Akavia, Goldwasser, and Vaikuntanathan~\cite{akavia2009} to show that $X$ has many hardcore bits.  The result says that if $\distLWE_{(n-k, m, q, \chi)}$ is secure then any $k$ variables of $X$ in a $\distLWE_{(n, m, q, \chi)}$ instance are hardcore.  We state their result for a general error distribution~(noting their proof does not consider the error distribution anywhere):
\begin{lemma}\protect{\cite[Lemma 2]{akavia2009}}
\label{lem:many hardcore bits}
Let $n>0$ be an integer, $q\geq 2$, let $\chi$ be an error distribution over $\zq$ and $k\leq n$ be an integer.  Let $\vA\in \zq^{m\times n} , X\in\zq^n$ and $E$ be distributed according to $\chi^m$.  If $\distLWE_{(n-k, m, q, \chi)}$ is $\epsilon, s_{sec}$ hard then $H^{\hill}_{\epsilon, s_{sec}}(X_{1,..., k}| \vA, \vA X+E )\geq k\log q$.
\end{lemma}
\bnote{Need to check on gap in the reduction, the claim here is that there is no loss.  I think it will go from $\ngl, \poly$ to $\ngl', \poly'$}

@@

\section{\consref{cons:informal construction} with Symbol Fixing Sources}
\label{sec:LWE block fixing sources}
In this section, we show the security of~\consref{cons:informal construction} is preserved for certain high entropy distributions.  Showing security of a computational fuzzy extractor for arbitrary high min-entropy distributions is an open problem.  First we recall the notion of a symbol fixing source~(from~\cite{KZ07}): 
\begin{definition}
Let $W = (W_1,..., W_{m+\alpha})$ be a distribution where $W_i$ takes values over $\{0,1\}^b$, we say that a distribution is a $m$-\emph{symbol fixing source} if $H_\infty(W)\geq mb$ and for all $W_i$ one of the following conditions is met:
\begin{itemize}
\item $W_i \overset{d}= U_b$
\item There exists a value $c_i$ such that $\Pr[W_i = c_i] =1$.
\end{itemize}
\end{definition}
%A block fixing source is an extension of a entropy source where each bit is independently random or fixed to larger alphabets.

@@

\begin{theorem}
\label{thm:lossless block sketch log}
Let $n$ be a security parameter and let $t = c\log n$ for some constant $c$.  Let $d\in\Fq^+$ be a constant and consider the Hamming metric with block length $b = \log (2n^3 (c+d)/d)$.  Let $W$ be $\alpha$-symbol fixing source over $\zo^{((c+d)nb/d+\alpha b}$.  Assume GAPSVP and SIVP are hard to approximate within polynomial factors~(on lattices of dimension $n$) using quantum algorithms.   There is a setting of $q = \poly(n)$ such that \consref{cons:informal construction} is a lossless fuzzy extractor.  That is, $H^{\hill}_{\epsilon, s_{sec}}(X_{1,..., n/2} |\vA, \vA X+W) = H_\infty(W)$ for %the following parameters:
%\[
%(\zo^{((c+d)n/d+\alpha)b}, \Hoo(W),\Hoo(W), \epsilon, s_{sec}, s_{sketch}, s_{rec}, t)\text{-computational secure sketch}
%\]
$\epsilon = \ngl(n), s_{sec} = \poly(n)$, sketch time $O(n^3\log n)$, and expected recover time $O(n^{2d+3} \log n)$.
\end{theorem}

@@

\begin{claim}
\label{cl:randomMatrixDist}
The matrix $\vA$ generated after Step \ref{step:randomization} is a random matrix.
\end{claim}
\begin{proof}
In the randomized matrix $\vect{A}$, for $1\le i \le m$ and $n+1\le j \le n+\alpha$, $A_{ij}$ is assigned the value $\gamma_{i,(j - n)}$, which is a random value.  Similarly, for $m+1\le i \le m+\alpha$ and $n+1\le j \le n+\alpha$, each entry $A_{ij}$ is random due to the original construction of the matrix $\vA^*$.  Then before step \ref{step:randomization} the matrices $\vA', \vect{R}$ are random submatrices.  Thus, it remains to show that $\vA', \vect{R}$ remain random after step 5. For $1\le i \le m+\alpha$ and $1\le j \le n$, by construction $A_{ij} = A_{ij}'+\sum_{k=1}^\alpha \gamma_{i, k} R_{kj}$, which is a truly random number. Overall, each row vector $\va_{i}$ in $\vA$ is the sum of a random vector and a random linear combination independent vectors. Therefore, the entire matrix $\vA$ is a truly random matrix.  This completes the claim.
\end{proof}
\begin{claim}
\label{cl:random ax+e}
If $\D'$ is provided with input distributed as $\vA', \vA'\vx'+\ve'$ then $\vb = \vA \vx+\ve$ where $x_i = x_i'$ for $1\leq i \leq n$ and $x_{n+i} = b_{m+i}^*$ is uniformly generated otherwise and  $e_i = e_i'$ for $1\leq i\leq m$ and $e_i = 0$ for $m<i\leq m+\alpha$.
\end{claim}
\begin{proof}
Before step \ref{step:randomization} the claim is satisfied.  Denote by $\vx'$ the value such that $\vA'\vx'+\ve = \vb$.  
Then define $\vx^*$ as:
\[
x^*_i = \begin{cases}
x_i', & \text{if }1\leq i\leq n \\
b_{m-n+i}^*-\sum_{j=1}^n R_{i, j} x_j', & n+1\leq i\leq n+\alpha\,.
\end{cases}
\]
Then $\vx^*$ is uniformly random and $\vb^* = \vA^*\vx^* + \ve^*$~(where $\ve^*$ is $\ve'$ with $\alpha$ 0s appended).  Thus, it remains to show that step~\ref{step:randomization} preserves a solution.
%For convenience, denote by $\vx^*$ the vector where $x_i^* = x_i'$ for $1\leq i \leq n$ and $x_{n+i}^* = b_{m+i}^*$ otherwise.  
For a matrix $\vA$ we will denote the $i$-th row by $\va_i$~(similarly for the matrix formed by $\gamma_{i,j}$).  Thus it suffices to show for all rows $i$, if $b_i^* = \va^*_i \vx^*+e^*_i$  then $b_i = \va_i \vx^* + e^*_i$.
%, we have added $\alpha$ equations as well as $\alpha$ unknowns, furthermore, those $\alpha$ equations have no error.  The number of unknowns we added are $x_{n+i} = b^*_{m+i} - \sum^{n}_{j=1}R_{ij}x_j$ for $i=n+1,..., n+\alpha$.
\bnote{We need to justify these steps better.}
\xnote{lower case for vectors, which is row vector in the matrix}

We have the following for all $i=1,..., m+\alpha$:
\begin{align*}
\va_i \vx^* + e_i^* &= \sum_{j=1}^{n+\alpha} \vA_{ij} x^*_j + e_i^*\\
& = \sum_{j=1}^{n+\alpha}\left((\vA^*_{ij}+\sum_{k=1}^\alpha \gamma_{i, k} \vA^*_{(m+k)j})x^*_j\right) + e_i^*\\
& = \sum_{j=1}^{n+\alpha}\vA^*_{ij}x^*_j + e_i^* + \sum_{j=1}^{n+\alpha}\sum_{k=1}^\alpha \gamma_{i, k} \vA^*_{(m+k)j}x^*_j \\
& = \sum_{j=1}^{n+\alpha}\vA^*_{ij}x^*_j + e_i^* + \sum_{k=1}^\alpha  \gamma_{i, k} \sum_{j=1}^{n+\alpha} \vA^*_{(m+k)j}x^*_j\\
& = \sum_{j=1}^{n+\alpha}\vA^*_{ij}x^*_j + e_i^* + \sum_{k=1}^\alpha \gamma_{i, k} \left(\left( \sum_{j=1}^n \vR_{i, j}x_j^* \right) + x_{n+k}^*\right)\\
& = \sum_{j=1}^{n+\alpha}\vA^*_{ij}x^*_j + e_i^* + \sum_{k=1}^\alpha\gamma_{i, k} b_{m+k}^*\\
&= b_i^* +\sum_{k=1}^\alpha\gamma_i b_{m+k}^*= b_i
\end{align*}
%\begin{align*}
%\vA_i \vx + e_i &= \sum_{j=1}^{n+\alpha} A_{ij} x_j +e_i = \sum_{j=1}^{n+\alpha}(A^*_{ij}x_j+\sum_{k=1}^\alpha \gamma_{i, k} A^*_{(m+k)j}x_j)+e_i\\
%& = \sum_{j=1}^{n+\alpha}(A^*_{ij}x_j)+e_i +\sum_{k=1}^\alpha \gamma_{i, k} \sum_{j=1}^{n+\alpha} (A^*_{(m+k)j}x_j)\\
%& = b^*_i +  \sum_{k=1}^\alpha\gamma_{i, k} b^*_{m+k} = (\va_i^* \vx + e_i) + \sum_{k=1}^\alpha\gamma_{i, k} (\va_{m+k}^* \vx + e_i)\\
%& = b_i \label{reductionEq}
%\end{align*}
\end{proof}
\begin{claim} 
\label{clm:random b}
If $\D'$ is provided with the input distributed as $\vA', \vb'$ where $\vb'$ is uniformly random, then $\vb$ is uniformly random conditioned on $\vA$.
\end{claim}
\begin{proof}
Before step~\ref{step:randomization}, each $b_i^*$ is uniformly and independently random.  After step~\ref{step:randomization}, since the $b_i*$ and $\gamma_i$ are random, then each $b_i$ is a fresh random linear combination of random values:
\[
b_i = b_i^* + \sum_{k=1}^\alpha \gamma_i b_{m+k}^*.
\]
%We also need to show that $\vb$ is uniformly distributed even conditioned on $\vA$. Step~\ref{step:randomization} randomized each

$\vb$ doesn't dependent on $\vA$ simply because $\vb'$ doesn't dependent on $\vA'$ \xnote{It need show more work on this, maybe?}
Thus, each of these values is uniformly and independently random.
\end{proof}
